The Etiology of Cybersecurity

### 2020-07-21 reports/paper_0/main.tex

Our hypothesis is that cybersecurity attacks are caused by 
the presence of errors in the design or implementation of a system.
For example, a design or implementation error in a sanitization function
may allow an attacker to perform SQL-injections. If errors are necessary
to enable cybersecurity attacks, a theory that predicts all potential 
errors (or source of errors) can be effectively used to predict the
security of a system. With our approach, a list of weaknesses emerges
from the mathematical formulation of the system and predicts 4 main classes
of weaknesses. Those classes of weaknesses are used to calculate
all the insecurity configurations of all the components of a system,
obtaining a precise estimation of all potential cybersecurity-related risks
in any given system.

@@

Currently, the most complete understanding of insecurity issues, is stored into
a network of databases of weaknesses (e.g.
CWE~\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE~\autocite{CVE},
NVD~\autocite{NIST2020NVD}), and attacks (e.g. CAPEC~\autocite{MITRE2020CAPEC},
ATT\&CK~\autocite{MITRE2020ATTACK}).  Those insecurity issues can be related to
the violation of one or more requirements (explicit or implicit) in the
specification, design or implementation of a system. The correlation 
between insecurity flaws and security requirements has been used to define
standards such as the IEC 62443-1-3 (the Industrial communication networks -
Network and system security -- Part 3-3: System security requirements and
security levels) defines requirements as ``confidentiality of information
in transit/at-rest''. More generally, the idea of defining security requirements
as properties of a system was initially defined in 1970s with the 
CIA-triad (Confidentiality, Integrity, Availability) and refined 
over the decades introducing related concepts such as authenticity or
non-repudiation, or introducing new ones such as ``responsibility'' in the
RITE approach (see~\autocite{Samonas2014cia} for an overview of the evolution of the CIA-triad).

Systems are
nowadays tested against known or unknown attacks,
or verified against security requirements such as confidentiality.
However, no scientific falsifiable theory correlate
security requirements to necessary and sufficient conditions (e.g. mitigations) 
to declare a system secure.

\subsection{Terminology}
Most of the safety-preserving principles in the field of engineering of
safety-critical cyber-physical systems (such as elevators and aircraft), upon
which safety requirements are defined (e.g. in standards such as the IEC 61508
or 61511\autocite{IEC201761511}), are based on empirical tests and measurements
(therefore should be considered hypotheses and not definitions).\fixnote{Luca}{Hypotheses and definitions don't type check. Moreover, many standards require formal methods proofs, don't they?} While
reasoning by induction
%\footnote{``So, whenever they argue ``Every man is an
%animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
%deduce from the universal proposition ``every man is an animal'' the particular
%proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
%mentioned) to establish by way of induction the universal proposition, the
%fall into the error of circular reasoning, since they are establishing the
%universal proposition inductively by means of each of the particulars and
%deducing the particular proposition from the universal syllogistically.''
%Sextus Empiricus, Outlines of Pyrrhonism II-195 \autocite{Empiricus1990Pyrrhonism}}
based on the empirical observation should be avoided, since it may easily lead
to false beliefs, this approach is often justified by the supposed
impossibility of defining a theory that correctly predicts failures.\fixnote{Luca}{This sentence does not follow grammatically}  A failure
of a wire due to environment (e.g. due to humidity, dust, heat \&c) is defined
from empirical evidences and processes have been standardized to test qualities
of hardware components.  This process completely breaks down when a malicious
environment (i.e. an attacker) is considered instead of the (supposedly honest
and predictable) natural environment. Therefore, the same approach that is in
use for safety, seems not to be applicable for security (e.g. for security
testing).
An overview on the aforementioned aspects of safety and security is depicted in
Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
the terms that structure our current understanding of security.\fixnote{Luca}{I disagree. It should be cybersecurity, not security. Moreover, as I mentioned, the figure is misleading when it talks about blac-hat hacking. I also don't like ``Manking'', as mankind does many of the things, not just designing systems} 
\begin{itemize}
	\item \emph{Mankind} ``refers collectively to humans''
\autocite{wiki-mankind}, while the concept of \emph{Nature} is
		related ``to the intrinsic characteristics that plants,
		animals, and other features of the world develop of their own
		accord'' (e.g. the physical universe)~\autocite{wiki-nature}. 
		\begin{itemize}
			\item There exist several terms to refer to an
				\emph{attacker}, i.e. threat agent or threat source,
				considering those terms to be semantically
				equivalent.\fixnote{Luca}{Grammar issues}  From now on, we consider the
				Causality principle to be the \emph{threat
				source}, Nature or Mankind to be the
				\emph{threat agents} and an \emph{attacker} as
				a specific malicious threat agent which materializes a
				threat.
	\item \emph{Vulnerability}\footnote{The term vulnerability is not
		present in the Encyclopedia of Cryptography and Security, while
		it is used in 12 entries (such as in the definition of
		``penetration testing'' \autocite{caddy2005pentest})
		highlighting how commonly this word is used without a proper
		supporting semantics.}, as defined in \autocite{cnssi20104009}
		(and adopted in \autocite{nist2013800-53}), is a ``weakness in an
		information system, system security procedures, internal
		controls, or implementation that could be exploited by a threat
		source''. On the one hand, the definition is broad to enclose
		as much causes (that generates a vulnerability) as possible,
		on the other hand the term vulnerability should have a complete and sound
		definition, so that no other causes (e.g.  other sources) but
		the ones in the definition are responsible for a vulnerability
		%\footnote{``For this view, that
		%\emph{That Which Is Not} exists, can never predominate. You
		%must debar your thought from this way of search, nor let
		%ordinary experience in its variety force you along this way,
		%(namely, that of allowing) the eye, sightless as it is, and the
		%ear, full of sound, and the tongue, to rule; but (you must)
		%judge by means of the Reason (Logos) the much-contested proof
		%which is expounded by me.'' -- Parmenides of Elea, On Nature
		%(circa 500 B.C.), fragments B7.1–8.2
		%\autocite{Hakim2016philosophy}}.
		Furthermore, the term ``threat sources'' used in the definition
		in~\autocite{cnssi20104009} may be identified with both Nature
		and Mankind, not differentiating between safety and security.

@@

\begin{itemize}
	\item \emph{Weakness}. The definition given by the MITRE
		in \autocite{MITRE2020CWEweakness} of weakness is: ``Software
		weaknesses are errors that can lead to software
		vulnerabilities.\fixnote{Luca}{Worth to say that many security issues are not software weakenesses, e.g. cryptography issues or the way cryptography is used. Moreover, in cyber-physical systems, security issues might be due to hardware issues. Again, I think that you are trying to say too many things. Focus on what you really want to say.} A software vulnerability, such as those
		enumerated on the Common Vulnerabilities and Exposures (CVE)
		List, is a mistake in software that can be directly used by a
		hacker to gain access to a system or network''.  The definition
		is circular if we interpret the word ``error'' and ``mistake''
		with the same semantics: a weakness is an error that leads to a
		vulnerability and a vulnerability is a mistake which, in turn,
		is a weakness.  
\end{itemize}

The only difference between a weakness and vulnerability seems to be that one
can consider weakness as a ground term and state that a vulnerability is caused
by a weakness.

\begin{itemize}
	\item \emph{Causality} refers to the causality principle; defined
		in~\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
		connection of phenomena through which one thing (the cause)
		under certain conditions gives rise to, causes something else
		(the effect). The essence of causality is the generation and
		determination of one phenomenon by another. In this respect,
		causality differs from various other kinds of connection, for
		example, the simple temporal sequence of phenomena, of the
		regularities of accompanying processes''. 
		%One can consider the
		%causality principle to be formalized by the K Modal Logic.
	\item An \emph{Exploit}\footnote{We note that the term exploit is only
		used as a verb in~\autocite{ISO2009information}}\fixnote{Luca}{So what? Is it The use in~\autocite{ISO2009information} is incomplete? } ``[\ldots]
		(from the English verb to exploit, meaning to use something to
		one’s own advantage) is a piece of software, a chunk of data,
		or a sequence of commands that takes advantage of a bug or
		vulnerability to cause unintended or unanticipated behavior to
		occur on computer software, hardware, or something electronic
		(usually computerized).''~\autocite{wiki-exploit}.
	\item An \emph{Attack}, as defined by the International Standard
		ISO/IEC 27000, is an ``attempt to destroy, expose, alter,
		disable, steal or gain unauthorized access to or make
		unauthorized use of an asset''; where an \emph{Asset} is
		``anything that has value to the organization''. We note that for
		the purpose of this article, we do not want to focus on a specific
		organization or business to define asset but, in general, on any 
		abstract organization (e.g. a company or a society).
		We do not consider ethical hackers as attacking a system. 
		In fact, we consider the term \emph{hack} as
		non-malicious (as, e.g. in~\autocite{Stallman2002hacker}).
	\item A \emph{Threat}, as defined in~\autocite{cnssi20104009}, is ``Any
		circumstance or event with the potential to adversely impact
		organizational operations (including mission, functions, image,
		or reputation), organizational assets, individuals, other
		organizations, or the Nation through an information system via
		unauthorized access, destruction, disclosure, modification of
		information, and/or denial of service''.
	\item \emph{Defect}: ``anything that renders the product not reasonably
		safe''~\autocite{Robinson2019writing} (i.e. a characteristic of
		an object which hinders its proper usability).
	\item \emph{Failure}, as defined in\autocite{Merriam2020failure}, is ``a state of
		inability to perform a normal function''. The term is
		structured and detailed in
		\autocite{cnssi20104009,iet2017glossary}, but relying on an
		abstract notion of failure without a specific definition.
	\item \emph{Hazard}, ``a potential source of
		harm''~\autocite{iet2017glossary}.
\end{itemize}

@@

Our literature review shows that most of the definitions relate insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker).\fixnote{Luca}{This is slightly misleading, as you have not discussed this above. You mean in the review that you did you discovered that...} This, however, just moves the problem of
defining what cybersecurity is as the problem of defining what a dishonest
agent can or cannot do in a system, where an agent is any virtual or
physical entity of the system or using the system (e.g. a device, a software,
or a human being) and dishonesty is not necessarily related to malicious
motivation but also to incompetence or lack of skills. As in the Dolev-Yao
theory, we may correlate ``being dishonest'' to ``not following the intended
behavior/rules''.  In the case of a generic system, a dishonest agent is,
therefore, any agent that doesn't follow the intended behavior (or
functionality or logic) of the system.  Given the generality of the definition,
and its high level of abstraction, we may conclude that the hypothesis seems
evident.\fixnote{Luca}{Which hypothesis?} For example, a software can be considered an agent of the system and
whenever it has a bug, it can be exploited causing an Incident. However, this
hypothesis has lead to the un-testable conclusion that the dishonest behaviors
of agents cannot be defined in general (e.g. due to the heterogeneity of agents
and systems) and that huge repository of dishonest behaviors should be kept as
definition, such as CAPEC~\autocite{MITRE2020CAPEC}; or that dishonesty of an agent
with respect to a security protocol should be defined as a number of predefined
actions as in
\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few).\fixnote{Luca}{I foudn this sentence very hard to follow. Moreover, I would be careful to speak of ``predefined actions'' as basically the DY can do anything so it is not cleaer what you are referring to. Of course, you could this that this excludes, e.g., human errors (say as in the work I did with Diego) or more, but then you should clarify.} As depicted in Figure~\ref{fig:causality}, in order to define
a theory on cybersecurity and on the malicious behaviors of agents we noticed
that Attack Patterns (or Attacks) are caused-by the presence of Vulnerabilities
in the system.\fixnote{``cause-by'' is grammatically wrong, but more importantly the presence of a vulnerability does not necessarily cause an attack, but allows it. Moreover, you speak of Patterns for the first time.} Those Vulnerabilities are, in turn, cause-by the presence of
Weaknesses in the system.  Weaknesses are Errors in the design or
implementation of a system.  Therefore, a theory on cybersecurity should first
predict the Errors in a system design.

@@

Similarly to\autocite{Santaca2016abf}, we define agents as a meronomy (an
hierarchy of Part-Whole relations) over the constituent that we previously defined
(assertions, beliefs, and facts), based on a standard definition of mereology,
i.e. based on the definition of Parthood relation between \emph{Parts}.  Due to
the necessity of considering different types of Part (as we'll show afterwards)
we extend the mereology to a
mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Table~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{Region} both to refer to a mereological Part
and to a topological Region.  Our aim is to create a meronomy (hierarchy of part-whole
relations)
instead of the taxonomies (categorization based on discrete sets) 
such as the one provided
in\autocite{NIST2020NVD,MITRE2020CVE} 
so that we don't need to rely on a scoring system (such as the 
CVSS\autocite{Mell2007CVSS}) to assign a quantitative evaluation 
of the security of each entry. Instead, we want a precise calculation
of the number of insecurity configurations to emerge from the mathematical
formulation of our cybersecurity theory.

@@

\paragraph{Security Weaknesses -- The RIDI-Hypothesis} 
It is important to note that all the results of the application of the
$\abf$-framework to channels (the analysis of the relation
$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) lead to the same results of the
analysis of a pair of an input and output port (i.e.
$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
$\Rcc{\assertionRegion_2}{\beliefRegion_3}$).  The same result can be obtained
by analyzing the relation between the outputs of a functional block and the
inputs of another functional block ($\Rcc{\beliefRegion_o}{\beliefRegion_i}$), 
where functional blocks are constituents of
the functional architecture as described afterwards.  As depicted in
Figure~\ref{fig:atom}, so far we have considered information generated by a
port $P_I$ and then sent through a channel $C$ to another (recipient) port
$P_O$. In this scenario where ports and then channels are atomic (otherwise
raising infinite regress), we can only  consider the relations between ports
and channel; considering both input-port to channel and channel to output-port.
In fact, the Weaknesses of a channel are defined in terms of the Weaknesses of
ports.  Io order to define a functional block
without encountering a infinitely recursive definition, we must reach the same
conclusions as for the channel. Therefore, describing the information as
flowing over a channel or in a functional block is purely syntactic sugar.

>>>>

Instead of reasoning of what makes a system secure or insecure,
we reason on what causes insecurities.
Our hypothesis is that cybersecurity attacks are only caused by 
the presence of errors in the design or implementation of a system.
For example, a design or implementation error in a sanitization function
may allow an attacker to perform SQL-injections. If errors are necessary
to enable cybersecurity attacks, a theory that predicts all potential 
errors (or source of errors) can be effectively used to predict the
security of a system. With our approach, a list of weaknesses emerges
from the mathematical formulation of the system and predicts 4 main classes
of weaknesses. Those classes of weaknesses are used to calculate
all the insecurity configurations of all the components of a system,
obtaining a precise estimation of all potential cybersecurity-related risks
in any given system. Our hypothesis can be falsified by means of experiments,
testing if all the predicted weaknesses are present in the system under
test, or testing if other (not predicted by our hypothesis) weaknesses are
present.

@@

Currently, the most complete understanding of insecurity issues, is stored into
a network of databases of weaknesses (e.g.
CWE~\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE~\autocite{CVE},
NVD~\autocite{NIST2020NVD}), and attacks (e.g. CAPEC~\autocite{MITRE2020CAPEC},
ATT\&CK~\autocite{MITRE2020ATTACK}).  Those insecurity issues can be related to
the violation of one or more requirements (explicit or implicit) in the
specification, design or implementation of a system. The correlation 
between insecurity flaws and security requirements has been used to define
standards such as the IEC 62443-1-3 (the Industrial communication networks -
Network and system security -- Part 3-3: System security requirements and
security levels) which defines requirements as ``confidentiality of information
in transit/at-rest''. More generally, the idea of defining security requirements
as properties of a system was initially defined in 1970s with the 
CIA triad (Confidentiality, Integrity, Availability) and refined 
over the decades introducing related concepts such as authenticity or
non-repudiation, or introducing new ones such as ``responsibility'' in the
RITE approach (see~\autocite{Samonas2014cia} for an overview of the evolution of the CIA triad).
The link between security requirements and vulnerabilities is reported 
in the NVD databases by the CVSS\autocite{Mell2007CVSS}) scoring system.
The CVSS evaluates of the severity of a vulnerability
by means of different metrics (such as attack complexity and user interaction)
and quantitatively evaluates the impact on the CIA triad.
While security requirements, weaknesses, vulnerabilities, and attacks
have been extensively studied and implemented both in academia and industry
to provide tools for the testing or verification of systems, 
no scientific falsifiable theory correlate
security requirements to necessary and sufficient conditions (e.g. mitigations) 
to declare a system secure \autocite{Herley2016unfalsifiability}.
Nonetheless, the extensive body of literature have scientific foundations,
for example, providing either providing necessary or sufficient conditions
for cybersecurity.
As a driver for our argumentation, we start by reviewing the key concepts
in the cybersecurity domain.

\subsection{Terminology}
Most of the safety-preserving principles in the field of engineering of
safety-critical cyber-physical systems (such as elevators and aircraft), upon
which safety requirements are defined (e.g. in standards such as the IEC 61508
or 61511\autocite{IEC201761511}), are based on empirical tests and measurements.
While reasoning by induction
%\footnote{``So, whenever they argue ``Every man is an
%animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
%deduce from the universal proposition ``every man is an animal'' the particular
%proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
%mentioned) to establish by way of induction the universal proposition, the
%fall into the error of circular reasoning, since they are establishing the
%universal proposition inductively by means of each of the particulars and
%deducing the particular proposition from the universal syllogistically.''
%Sextus Empiricus, Outlines of Pyrrhonism II-195 \autocite{Empiricus1990Pyrrhonism}}
based on the empirical observation should be avoided in general,
since it may easily lead to false beliefs \autocite{Popper1959logic}, 
inducing general principle by means of tests
is often justified by the supposed impossibility of defining a theory
that correctly predicts failures.
A failure of a wire due to environment (e.g. due to humidity,
dust, heat \&c) is defined from empirical evidences and processes have been
standardized to test qualities of hardware components.  This process completely
breaks down when a malicious environment (i.e. an attacker) is considered
instead of the (supposedly honest and predictable) natural environment.
Therefore, the same approach that is in use for safety, seems not to be
applicable for cybersecurity (e.g. for cybersecurity testing).  An overview on the
aforementioned aspects of safety and cybersecurity is depicted in
Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
the terms that structure our current understanding of cybersecurity.

\begin{itemize}
	\item \emph{Mankind} ``refers collectively to humans''
\autocite{wiki-mankind}, while the concept of \emph{Nature} is
		related ``to the intrinsic characteristics that plants,
		animals, and other features of the world develop of their own
		accord'' (e.g. the physical universe)~\autocite{wiki-nature}. 
		\begin{itemize}
			\item There exist several terms to refer to an
				\emph{attacker}, e.g. threat agent or threat source,
				but, from now on, we consider the
				Causality principle to be the \emph{threat
				source}, Nature or Mankind to be the
				\emph{threat agents} and an \emph{attacker} as
				a specific malicious threat agent which materializes a
				threat.
		\end{itemize}
	\item \emph{Vulnerability}\footnote{The term vulnerability is not
		present in the Encyclopedia of Cryptography and Security, while
		it is used in 12 entries (such as in the definition of
		``penetration testing'' \autocite{caddy2005pentest})
		highlighting how commonly this word is used without a proper
		supporting semantics.}, as defined in \autocite{cnssi20104009}
		(and adopted in \autocite{nist2013800-53}), is a ``weakness in an
		information system, system security procedures, internal
		controls, or implementation that could be exploited by a threat
		source''. On the one hand, the definition is broad to enclose
		as much causes (that generates a vulnerability) as possible,
		on the other hand the term vulnerability should have a complete and sound
		definition, so that no other causes (e.g.  other sources) but
		the ones in the definition are responsible for a vulnerability
		%\footnote{``For this view, that
		%\emph{That Which Is Not} exists, can never predominate. You
		%must debar your thought from this way of search, nor let
		%ordinary experience in its variety force you along this way,
		%(namely, that of allowing) the eye, sightless as it is, and the
		%ear, full of sound, and the tongue, to rule; but (you must)
		%judge by means of the Reason (Logos) the much-contested proof
		%which is expounded by me.'' -- Parmenides of Elea, On Nature
		%(circa 500 B.C.), fragments B7.1–8.2
		%\autocite{Hakim2016philosophy}}.
		Furthermore, the term ``threat sources'' used in the definition
		in~\autocite{cnssi20104009} may be identified with both Nature
		and Mankind, not differentiating between safety and security.
\end{itemize}

@@

\begin{itemize}
	\item \emph{Weakness}. The definition given by the MITRE in
		\autocite{MITRE2020CWEweakness} of weakness is: `` a type of
		mistake that, in proper conditions, could contribute to the
		introduction of vulnerabilities within that product. This term
		applies to mistakes regardless of whether they occur in
		implementation, design, or other phases of a product
		lifecycle.'' A vulnerability, such as those enumerated on the
		Common vulnerabilities and Exposures (CVE) List, is a mistake
		that can be directly used by a hacker to gain access to a
		system or network.  The definition is circular if we interpret
		the word ``error'' and ``mistake'' with the same semantics: a
		weakness is an error that leads to a vulnerability and a
		vulnerability is a mistake which, in turn, is a weakness. The
		only difference between a weakness and vulnerability seems to
		be that one can consider weakness as a ground term and state
		that a vulnerability is caused by a weakness.
	\item \emph{Causality} refers to the causality principle; defined
		in~\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
		connection of phenomena through which one thing (the cause)
		under certain conditions gives rise to, causes something else
		(the effect). The essence of causality is the generation and
		determination of one phenomenon by another. In this respect,
		causality differs from various other kinds of connection, for
		example, the simple temporal sequence of phenomena, of the
		regularities of accompanying processes''. 
		%One can consider the
		%causality principle to be formalized by the K Modal Logic.
	\item An \emph{Exploit} ``[\ldots]
		(from the English verb to exploit, meaning to use something to
		one’s own advantage) is a piece of software, a chunk of data,
		or a sequence of commands that takes advantage of a bug or
		vulnerability to cause unintended or unanticipated behavior to
		occur on computer software, hardware, or something electronic
		(usually computerized).''~\autocite{wiki-exploit}.
	\item An \emph{Attack}, as defined by the International Standard
		ISO/IEC 27000, is an ``attempt to destroy, expose, alter,
		disable, steal or gain unauthorized access to or make
		unauthorized use of an asset''; where an \emph{Asset} is
		``anything that has value to the organization''. We note that for
		the purpose of this article, we do not want to focus on a specific
		organization or business to define asset but, in general, on any 
		abstract organization (e.g. a company or a society).
		We do not consider ethical hackers as attacking a system. 
		In fact, we consider the term \emph{hack} as
		non-malicious (as, e.g. in~\autocite{Stallman2002hacker}).
	\item A \emph{Threat}, as defined in~\autocite{cnssi20104009}, is ``Any
		circumstance or event with the potential to adversely impact
		organizational operations (including mission, functions, image,
		or reputation), organizational assets, individuals, other
		organizations, or the Nation through an information system via
		unauthorized access, destruction, disclosure, modification of
		information, and/or denial of service''.
	\item \emph{Defect}: ``anything that renders the product not reasonably
		safe''~\autocite{Robinson2019writing} (i.e. a characteristic of
		an object which hinders its proper usability).
	\item \emph{Failure}, as defined in\autocite{Merriam2020failure}, is ``a state of
		inability to perform a normal function''. The term is
		structured and detailed in
		\autocite{cnssi20104009,iet2017glossary}, but relying on an
		abstract notion of failure without a specific definition.
	\item \emph{Hazard}, ``a potential source of
		harm''~\autocite{iet2017glossary}.
\end{itemize}

@@

Our literature review shows that most of the definitions relate insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker). In fact, weaknesses become vulnerability if an
attacker exploits them in an attack.  This, however, just moves the problem of
defining what cybersecurity is as the problem of defining what a dishonest
agent can or cannot do in a system, where an agent is any virtual or physical
entity of the system or using the system (e.g. a device, a software, or a human
being) and dishonesty is not necessarily related to malicious motivation but
also to incompetence or lack of skills. As in the Dolev-Yao theory, we may
correlate ``being dishonest'' to ``not following the intended behavior/rules''.
In the case of a generic system, a dishonest agent is, therefore, any agent
that doesn't follow the intended behavior (or functionality or logic) of the
system.  For example, a software can be considered an agent of the system and
whenever it has a bug, it can be exploited causing an Incident. However, this
may lead to the conclusion that the dishonest behaviors of agents cannot be
defined in general (e.g. due to the heterogeneity of agents and systems) and
that huge repository of dishonest behaviors should be kept as definition, such
as CAPEC~\autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
to a security protocol should be defined as a number of predefined actions as
in
\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few)\footnote{In formal protocol verifiers, in the DY model, the
attacker has a set of hardcoded inference rules that defines how the attacker
deduces new messages based on the observed message exchange.}.  As depicted in
Figure~\ref{fig:causality}, in order to define a theory on cybersecurity we may
say that the presence of vulnerabilities is a necessary condition in order to
cause an Attack in the system.  Those vulnerabilities are, in turn, caused by
the presence of weaknsesses in the system.  Weaknesses are errors in the design
or implementation of a system.  Therefore, a theory on cybersecurity should
first predict the errors in a system design.

@@

Similarly to\autocite{Santaca2016abf}, we define agents as a meronomy (an
hierarchy of Part-Whole relations) over the constituent that we previously defined
(assertions, beliefs, and facts), based on a standard definition of mereology,
i.e. based on the definition of Parthood relation between \emph{Parts}.  Due to
the necessity of considering different types of Part (as we'll show afterwards)
we extend the mereology to a
mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Table~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{Region} both to refer to a mereological Part
and to a topological Region.  Our aim is to create a meronomy (hierarchy of part-whole
relations)
instead of the taxonomies (categorization based on discrete sets) 
such as the one provided
in\autocite{NIST2020NVD,MITRE2020CVE} 
so that we don't need to rely on a scoring system (such as the 
CVSS) to assign a quantitative evaluation 
of the security of each entry. Instead, we want a precise calculation
of the number of insecurity configurations to emerge from the mathematical
formulation of our cybersecurity theory.

@@

\paragraph{Security Weaknesses -- The RIDI-Hypothesis} 
It is important to note that all the results of the application of the
$\abf$-framework to channels (the analysis of the relation
$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) lead to the same results of the
analysis of a pair of an input and output port (i.e.
$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
$\Rcc{\assertionRegion_2}{\beliefRegion_3}$).  The same result can be obtained
by analyzing the relation between the outputs of a functional block and the
inputs of another functional block ($\Rcc{\beliefRegion_o}{\beliefRegion_i}$), 
where functional blocks are constituents of
the functional architecture as described afterwards.  As depicted in
Figure~\ref{fig:atom}, so far we have considered information generated by a
port $P_I$ and then sent through a channel $C$ to another (recipient) port
$P_O$. In this scenario where ports and then channels are atomic (otherwise
raising infinite regress), we can only  consider the relations between ports
and channel; considering both input-port to channel and channel to output-port.
In fact, the weaknesses of a channel are defined in terms of the weaknesses of
ports.  Io order to define a functional block
without encountering a infinitely recursive definition, we must reach the same
conclusions as for the channel. Therefore, describing the information as
flowing over a channel or in a functional block is purely syntactic sugar.
